Unveiling parlr

### 2021-03-24 notes/Notes_3_26_21.Rmd
plot with constant nB, sampling demonstration

- Enamorado uses the following hashing function to identify unique aggreement patterns

@@

@@

- Each agreement pattern corresponds to a particular type of contribution to the posteriors of the $m$ and $u$ parameters. 
- To update $m$, I used to sum over each variable for all $N$ record pairs. I simply identify the number of matchings in the $Z$ vector corresponding to each agreement of $P$ patterns.

@@

- Again, I used to do this for each of $N$ records, but now I calculate this for the $P$ unique patterns.

- Just these two changes sped up the code considerably. However, I thought of one other clever trick to really push this idea to the max.

@@

- Rather than sampling $Z_j$ from $N$ many options, I instead sample it from the $P$ unique patterns, each one multiplied by its prevalance for record $j$.

@@

@@

- Only major hurdle with $N$ is calculating the comparison vectors, which is just unavoidable in the Fellegi Sunter Framework

@@

@@

>>>>

- Enamorado uses the following hashing function to identify unique aggreement patterns. I've also figured out how to do this efficiently in `dplyr`.

@@

- Within a Gibbs iteration, all records with the same agreement pattern will have the same FS weight. So instead of calculating the weight for each of $n_A \times n_B$ records, I only need to do it for $P$ unique patterns.

- This step is taken directly from `fastlink`.

@@

- `fastlink` doesn't have posterior updates for Dirichlet distributions, so I came up with this step on my own.

- Each agreement pattern corresponds to a particular type of contribution to the posteriors of the $m$ and $u$ parameters.

- To update $m$, I used to sum over each variable for all $N$ record pairs. Now, I simply identify the number of matchings in the $Z$ vector corresponding to each of $P$ patterns.

@@

@@

- Rather than sampling $Z_j$ sequentially from $N$ many records, I can sample it in parallel from $P$ many equivalence classes of unique patterns, each one multiplied by its prevalance for record $j$.

@@

- Only major hurdle with $N$ is calculating the comparison vectors, which is just unavoidable in the Fellegi Sunter Framework. However, whatever Ted does for `fastlink` would work here too.

@@

@@

## More on Speed Up

- Within the Gibbs sampler, calculations only depend on $n_B$ (the number samples I need to take to fill a $Z$ vector), and $P$ (the number of unique patterns). $n_A$ only influences the scale of the counts for the unique patterns.

- $n_A$ only comes to play in forming the comparison vectors and in back-filling the samples of $Z$ at the end of the sampler.

- This suggests that the sampler should be *almost constant* in $n_A$ (ie, the Gibbs sampling for a $4000 \times 500$ problem should not be substantially longer than for a $500 \times 500$ problem)

- Particularly useful for linking a smaller dataset to a much larger dataset. One example would be the DNC data in BRACS!

## More on Speed Up

```{r}
knitr::include_graphics("figures/speed_plot_fixed_nB.png")
```

@@

## Limitations

- The original FS method is $O(n_A \times n_B)$ hard. With hashing methods, `fastlink` makes the problem only $O(P)$ hard, since record pairs are considered indepenently and classification is based *only* on the unique pattern.

- Hashing methods in my model make the problem $O(n_B \times P)$ hard. No matter what, I need to sample $n_B$ many times to fill a linkage vector $Z$. In this sense, `fastlink` is inherently faster than this method

- However, method takes into consider much more of the dependency structure in the data, making this computational burden worthwhile

- \textcolor{blue}{Note:} If this line of thinking is useful for the paper, I'll need some help making this rigorous!
